---
title: "Replication of College Football Games and Crime"
author: "Group 6"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```

# Introduction

This project replicates and extends Rees and Schnepel's study on whether college football games are associated with increases in crime in host communities. We focus on daily assault and vandalism counts from NIBRS and match them to college football schedules from 2000 to 2005.

# Data

*Placeholder.*

```{r load-data}
```

# Descriptive Statistics

## Table 1: Appendix

``` {r appendix-table}
analysis_data <- read_csv(
  "data/processed/analysis_data.csv",
  show_col_types = FALSE
)

team_agency <- read_csv(
  "data/processed/team_agency_mapping.csv",
  show_col_types = FALSE
)

analysis_data <- analysis_data %>%
  filter(ori %in% team_agency$ori)

appendix_table1 <- analysis_data %>%
  mutate(
    game_type = case_when(
      home_game == 1 ~ "Home Game",
      away_game == 1 ~ "Away Game",
      TRUE ~ "No Game"
    )
  ) %>%
  pivot_longer(
    cols = c(assault, vandalism),
    names_to = "crime_type",
    values_to = "count"
  ) %>%
  group_by(crime_type, game_type) %>%
  summarise(
    Mean = round(mean(count), 2),
    SD = round(sd(count), 2),
    Q25 = quantile(count, 0.25),
    Median = quantile(count, 0.50),
    Q75 = quantile(count, 0.75),
    Q90 = quantile(count, 0.90),
    .groups = "drop"
  )

knitr::kable(appendix_table1)
```

# Descriptive Statistics

## Table 2: Distribution of Game Days by Day of the Week

```{r table2}
table2 <- analysis_data %>%
  group_by(day_of_week) %>%
  summarise(
    Games = sum(home_game + away_game),
    Observations = n(),
    .groups = "drop"
  ) %>%
  mutate(
    day_of_week = factor(
      day_of_week,
      levels = c("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri")
    )
  ) %>%
  arrange(day_of_week)

table2 <- bind_rows(
  table2,
  tibble(
    day_of_week = "Total",
    Games = sum(table2$Games),
    Observations = sum(table2$Observations)
  )
)

knitr::kable(
  table2,
  col.names = c("Day of Week", "Games", "Observations")
)
```

# Descriptive Figures

*Placeholder.*

# Per-Team Plots

*Placeholder.*

# Regression Results

*Placeholder.*

# Extensions

*Placeholder.*

# Conclusion

*Placeholder.*