---
title: "Replication of College Football Games and Crime"
author: "Group 6"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```

# Introduction

This project replicates and extends Rees and Schnepel's study on whether college football games are associated with increases in crime in host communities. We focus on daily assault and vandalism counts from NIBRS and match them to college football schedules from 2000 to 2005.

# Descriptive Statistics

## Table 1: Appendix

```{r appendix-table}
analysis_data <- read_csv(
  "data/processed/analysis_data.csv",
  show_col_types = FALSE
)

analysis_data_clean <- analysis_data %>%
  filter(
    !team %in% c(
      "Ohio State",
      "Texas Tech",
      "South Carolina"
    )
  )

appendix_long <- analysis_data_clean %>%
  mutate(
    game_type = case_when(
      home_game == 1 ~ "Home Game",
      away_game == 1 ~ "Away Game",
      TRUE ~ "No Game"
    )
  ) %>%
  pivot_longer(
    cols = c(assault, vandalism),
    names_to = "crime_type",
    values_to = "count"
  )

appendix_table1 <- bind_rows(
  appendix_long %>% mutate(game_type = "All Days"),
  appendix_long
) %>%
  group_by(crime_type, game_type) %>%
  summarise(
    mean_sd = paste0(round(mean(count), 2), " (", round(sd(count), 2), ")"),
    q25 = as.character(quantile(count, 0.25)),
    q50 = as.character(quantile(count, 0.50)),
    q75 = as.character(quantile(count, 0.75)),
    q90 = as.character(quantile(count, 0.90)),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(mean_sd, q25, q50, q75, q90),
    names_to = "stat",
    values_to = "value"
  ) %>%
  mutate(
    stat = recode(
      stat,
      mean_sd = "",
      q25 = "0.25 quantile",
      q50 = "0.50 quantile",
      q75 = "0.75 quantile",
      q90 = "0.90 quantile"
    ),
    game_type = factor(
      game_type,
      levels = c("All Days", "No Game", "Home Game", "Away Game")
    ),
    crime_type = recode(
      crime_type,
      assault = "Assaults",
      vandalism = "Vandalism"
    )
  ) %>%
  arrange(crime_type, stat, game_type) %>%
  pivot_wider(
    names_from = game_type,
    values_from = value
  ) %>%
  select(
    crime_type,
    stat,
    `All Days`,
    `No Game`,
    `Home Game`,
    `Away Game`
  )

knitr::kable(
  appendix_table1,
  col.names = c("", "", "All Days", "No Game", "Home Game", "Away Game")
)
```

## Table 2: Distribution of Game Days by Day of the Week

```{r table2}
crime_daily <- read_csv("data/processed/crime_daily.csv", show_col_types = FALSE)
football_games <- read_csv("data/processed/football_games.csv", show_col_types = FALSE)
team_agency <- read_csv("data/processed/team_agency_mapping.csv", show_col_types = FALSE)

team_dates <- team_agency %>%
  crossing(date = seq(as.Date("2000-01-01"), as.Date("2005-12-31"), by = "day"))

home_games <- football_games %>%
  inner_join(team_agency, by = c("home_team" = "team")) %>%
  transmute(
    team = home_team,
    ori,
    date = game_date,
    home_game = 1,
    away_game = 0,
    home_win = home_win
  )

away_games <- football_games %>%
  inner_join(team_agency, by = c("away_team" = "team")) %>%
  transmute(
    team = away_team,
    ori,
    date = game_date,
    home_game = 0,
    away_game = 1,
    home_win = NA
  )

games_team_day <- bind_rows(home_games, away_games) %>%
  group_by(team, ori, date) %>%
  summarise(
    home_game = max(home_game),
    away_game = max(away_game),
    home_win = ifelse(any(home_game == 1), home_win[home_game == 1][1], NA),
    .groups = "drop"
  )

analysis_data <- team_dates %>%
  left_join(crime_daily, by = c("ori", "date" = "incident_date")) %>%
  left_join(games_team_day, by = c("team", "ori", "date")) %>%
  mutate(
    assault = replace_na(assault, 0),
    vandalism = replace_na(vandalism, 0),
    home_game = replace_na(home_game, 0),
    away_game = replace_na(away_game, 0),
    year = year(date),
    month = month(date),
    day_of_week = wday(date, label = TRUE)
  )

write_csv(analysis_data, "data/processed/analysis_data.csv")

# Replicate table 2

games_by_weekday <-
    read_csv("./data/processed/games_per_city_with_ori.csv") %>%
    mutate(weekday = wday(game_date, week_start = getOption("lubridate.week.start", 7), label = TRUE)) %>%
    group_by(weekday) %>%
    summarize(observation_days = n())

num_games <- 
    games_by_weekday %>%
    summarize(sum = sum(as.numeric(games_by_weekday$observation_days)))

game_location_oris <-
    read_csv("./data/processed/city_to_ori.csv") 

crimes_by_weekday <-
    read_csv("./data/processed/crime_daily.csv") %>%
    filter(ori %in% game_location_oris$ori, year(incident_date) >= 2000) %>%
    mutate(weekday = wday(incident_date, week_start = getOption("lubridate.week.start", 7), label = TRUE)) %>%
    group_by(weekday) %>%
    summarize(agency_days = n())

num_crime_days <- 
    crimes_by_weekday %>%
    summarize(sum = sum(as.numeric(crimes_by_weekday$agency_days)))


table2 <-
    left_join(crimes_by_weekday, games_by_weekday, by = "weekday") %>%
    add_row(weekday = "Total", agency_days = num_crime_days$sum, observation_days = num_games$sum) %>%
    replace_na(replace = list(observation_days = 0)) 
```

# Descriptive Figures

```{r descriptive-figures}
saturday_summary <- analysis_data %>%
  filter(day_of_week == "Sat") %>%
  mutate(
    game_type = case_when(
      home_game == 1 ~ "Home Game",
      away_game == 1 ~ "Away Game",
      TRUE ~ "No Game"
    )
  ) %>%
  group_by(game_type) %>%
  summarise(
    mean_assault = mean(assault),
    se_assault = sd(assault) / sqrt(n()),
    mean_vandalism = mean(vandalism),
    se_vandalism = sd(vandalism) / sqrt(n()),
    .groups = "drop"
  )

knitr::kable(saturday_summary)
```

```{r assault-figure}
ggplot(saturday_summary, aes(x = game_type, y = mean_assault)) +
  geom_col() +
  geom_errorbar(
    aes(
      ymin = mean_assault - se_assault,
      ymax = mean_assault + se_assault
    ),
    width = 0.2
  ) +
  labs(
    title = "Mean Saturday Assaults by Game Type",
    x = "Game Type",
    y = "Mean Assaults"
  )
```

```{r vandalism-figure}
ggplot(saturday_summary, aes(x = game_type, y = mean_vandalism)) +
  geom_col() +
  geom_errorbar(
    aes(
      ymin = mean_vandalism - se_vandalism,
      ymax = mean_vandalism + se_vandalism
    ),
    width = 0.2
  ) +
  labs(
    title = "Mean Saturday Vandalism by Game Type",
    x = "Game Type",
    y = "Mean Vandalism"
  )
```

# Per-Team Plots

```{r per-team-summary}
team_summary_home <- analysis_data %>%
  filter(day_of_week == "Sat") %>%
  mutate(
    game_type = case_when(
      home_game == 1 ~ "Home Game",
      home_game == 0 & away_game == 0 ~ "No Game",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(game_type)) %>%
  group_by(team, game_type) %>%
  summarise(
    mean_assault = mean(assault),
    mean_vandalism = mean(vandalism),
    .groups = "drop"
  ) %>%
  group_by(team) %>%
  filter(
    n_distinct(game_type) == 2,
    max(mean_assault, mean_vandalism) > 0
  ) %>%
  ungroup()
```

```{r per-team-assault}
ggplot(
  team_summary_home,
  aes(x = game_type, y = mean_assault, fill = game_type)
) +
  geom_col() +
  facet_wrap(~team, ncol = 4, scales = "free_y") +
  labs(
    title = "Mean Saturday Assaults: Home Game vs No Game",
    x = "Game Type",
    y = "Mean Assaults",
    fill = "Game Type"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )
```

```{r per-team-vandalism}
ggplot(
  filter(team_summary_home, team != "South Carolina"),
  aes(x = game_type, y = mean_vandalism, fill = game_type)
) +
  geom_col() +
  facet_wrap(~team, ncol = 4, scales = "free_y") +
  labs(
    title = "Mean Saturday Vandalism: Home Game vs No Game",
    x = "Game Type",
    y = "Mean Vandalism"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )
```

# Regression Results

*Placeholder.*

# Extensions

*Placeholder.*

# Conclusion

*Placeholder.*